\chapter{Introduction}

\begin{comment}

    Title
     - Fast launcher performance evaluation using reinforcement learning
     - Orbital ascent trajectory optimization using reinforcement learning
     - Orbital ascent trajectory optimization using reinforcement learning for the purpose of launcher performance evaluation
     - Fast launcher performance evaluation using reinforcement learning based guidance.
     - Fast orbital ascent trajectory optimization using reinforcement learning

    Describe problem(s) that we are trying to solve.
     - Launcher performance evaluation
     - Orbital ascent trajectory optimization
     - Slow performance of current launcher performance evaluation
     - Lifting body first stage
     - Unknown situations and circumstances

    
    Why it's relevant to the aerospace industry.
     - Space

    What the basic solution to this problem could be, current solutions
     - Global optimization methods

    
    Research question

    Proposed solution of the preliminary thesis and why.
     - RL
     - Short description of what RL is capable of
     - Give examples of previous research or real world application (keep short or out)

    Short description on the approach that will be taken during the thesis to use reinforcement learning to solve the problem.
     - Policy gradient
     - Self tuning quad-copter
     - Simulation

    Describe the contents in this report and how it will be used during the actual thesis.
     - Introduction into reinforcement learning
     - Tabular methods
     - Approximate solutions 
     - History of RL. Past, Present and Future (Or maybe just the latest in RL)
     - Thesis plan of attack


\end{comment}

Thanks to advancements in technology and an increasingly widespread use of satellite services, demand for space access has been steadily rising the past decades. This growth in the satellite launch services market has lead to the development of new launchers by both government agencies and an emerging private sector. 

One often overlooked, but crucial step during the design process of any product is the performance evaluation of proposed designs. This is the process of determining a products ability to meet the design requirements. Too low of a performance, and the product will not be able to fullfil its purpose. To high of a performance and the product may be over engineered or inefficient.

In the case of a launcher the performance may be measured by how much payload mass it can bring to it's target range of orbits. In other words, in order to maximize the performance of a launcher, the propellant mass fraction of the upper stage needs to be reduced.

% \todo{Maybe move this somewhere else. A bit too detailed for the introduction.}
% The typical flight profile of a launcher starts with a short vertical ascent followed by a pitch-over maneuver into the azimuth that will take the launcher to the desired orbital inclination. As the launcher slowly pitches down it increases both it's vertical and horizontal velocities, the former needed to reach the desired orbital altitude and the latter is required to successfully enter into a stable orbit. After a few minutes, the engine is cut-off and launcher continues coasting until is reaches it's apogee. Here the engine is ignited again in order to circularize the orbit and give it the desired eccentricity.

A successful orbital insertion requires the launcher to carefully follow an optimized trajectory. This trajectory has a massive effect on the performance of the launcher and is unfortunately not a trivial problem to solve. There has been quite some research in this trajectory optimization problem, but due to the complexity, the most reliable way to get the optimum is by using a global optimization method such as \acrfull{ga} or \acrfull{pso}.

Using a global optimizer to solve this problem is computationally intensive. This is not an issue when computing the trajectory for one particular launch during normal operations, but it is during the design the launch vehicle. A fast trajectory optimizer is desired during the design process in order to for example get a large number of results when performing a sensitivity analysis on the design parameters or continuously evaluate and monitor the feasibility of a changing design.

This thesis proposes the use of \acrfull{rl} as a possible solution to the trajectory optimization problem. With enough training an \acrshort{rl} agent is able to learn to solve complex and non-linear problems. The goal is to sufficiently train an \acrshort{rl} agent in order for it to be able to compute the optimal or close to optimal trajectory of a particular launcher design to a specific target orbit with just one simulation run. 

Even if generating optimal trajectories using \acrshort{rl} ends up being unfeasible, it may still generate close to optimal trajectories, which may either be acceptable for a design parameter sensitivity analysis, or serve as an initial guess for a more refined optimization algorithm.

The use of \gls{rl} in spacecraft guidance, navigation and control opens up the possibilities to solve future problems as space exploration becomes more prevalent. For example thanks to it's ability to run and adapt in real time, it may be used in situations that are difficult to forsee or too cumbersome to manually take into account, such as hovering and manoeuvering around the non-uniform, rotating gravitational field of an asteroid.

\emph{The objective of this thesis is to develop a orbital ascent trajectory optimization method using reinforcement learning and investigate its effectiveness compared to existing methods.}

\textbf{What is \acrlong{rl} and what can it be used for?}
\begin{itemize}
  \item How does \acrshort{rl} work?
  \item What is \acrshort{rl} capable of?
  \item How has \acrshort{rl} been used so far?
  \item What state of the art technological developments make use of \acrshort{rl}?  
\end{itemize}
  
\textbf{How are orbital ascent trajectories currently optimized?}
\begin{itemize}
  \item What are the challenges faced when optimizing an orbital ascent trajectory?
  \item What methods are currently used or being researched to solve this problem?
\end{itemize}

\textbf{How can \acrlong{rl} methods be used to solve the orbital ascent trajectory optimization problem?}
\begin{itemize}
  \item What advantages do \acrshort{rl} methods have over existing methods?
  \item What \acrshort{rl} methods have the potential of being effective in solving this problem?
  \item How do these newly developed methods compare to existing methods?
  \item May these \acrshort{rl} methods be used to compliment or improve the performance of existing methods?
\end{itemize}
