\chapter{Proposed thesis research}

\begin{comment}

    Describe problem that we are trying to solve.
     - complex dynamics of aerospace vehicles.
    
    Why it's relevant to the aerospace industry.
     - aerospace vehicle, duh, maybe this should go with the first item

    What the basic solution to this problem could be, current solutions
     - (adaptive) non-linear control

    Research questions. And what has been answered

    !! All of the above is also mentioned in the introduction !!
    Just write this information in the chapter introduction

    How RL could solve this problem

    The approach that will be taken to investigate this
     - Description of the environment
       • Quad-copter
         · States
         · Degrees of freedom
         · Controller description, gains
         · Which motion will the RL agent train for and what gains will it control, aka its actions
       • Rewards
         · Continuos rewards
         · Reward only at goal
         · Constant rewards while alive
       
     - What methods seem to be useful for the problem.
       • Value function only
         · (Expected) Sarsa
         · (Double) Q-learning
       • Policy Gradient
       • Some function or ANN
       • Eligibility traces?
       
     - How to test
      • Simulation
        · Custom setup around (P/C)ython, C/C++, JSBSim (or DARE FDM, this thing is experimental, but full freedom and knowledge to do whatever)
        · Parrot JSBSim simulation and Simulink setup used in previous thesis
      • Actual flight tests
        · What do we want to accomplish with these flight tests, why are they necessary.
        · States, motion and controller description is already in the environment section, maybe refer to that and give short summary.
        · Maybe collect info on hardware to get a sense of the computational limitations.
        
     - Expected results, requirements
      • What do we want the result to be in the end
       · Take-off?
       · Hover?
       · Move?
        - forwards
        - backwards
        - sideways
        - combination
       · Turn?
       · Follow flight path?
    
    Project expected timeline
     - What will happen in what order (Milestones)
     - Milestone deadlines?
\end{comment}